% ***************************************************
\section*{Publications included in this thesis}
As per UQ policy (\href{http://ppl.app.uq.edu.au/content/4.60.07-alternate-thesis-format-options}{PPL 4.60.07 Alternative Thesis Format Options}) the following section lists publications that have been included in this thesis.
\begin{enumerate}
    \item \textbf{Wilkinson, Ross D.}, Glen A. Lichtwark, and Andrew G. Cresswell. 2020. The Mechanics of Seated and Nonseated Cycling at Very-High-Power Output: A Joint-Level Analysis. \textit{Medicine $\&$ Science in Sports $\&$ Exercise} 52(7): 1584-94. doi: 10.1249/MSS.0000000000002285
    \item \textbf{Wilkinson, Ross D.}, Andrew G. Cresswell, and Glen A. Lichtwark. 2020. Riders Use Their Body Mass to Amplify Crank Power during Nonseated Ergometer Cycling. \textit{Medicine $\&$ Science in Sports $\&$ Exercise}: Publish Ahead of Print. doi: 10.1249/MSS.0000000000002408
\end{enumerate}

% ***************************************************
\section*{Submitted manuscripts included in this thesis}
\begin{enumerate}
    \item \textbf{Wilkinson, Ross D.}, Andrew G. Cresswell, and Glen A. Lichtwark. 2020. Rock and Roll: The Influence of Bicycle Lean on the Mechanics of Non-Seated Cycling. \textit{Journal of Biomechanics}: Under Review.
    \item \textbf{Wilkinson, Ross D.} and Glen A. Lichtwark. 2020. A Method for Tracking Centre Of Mass Displacement during Non-Seated Cycling using an Inertial Sensor. \textit{Journal of Biomechanics}: Under Review.
\end{enumerate}

% ***************************************************
\section*{Other publications during candidature}

% ***************************************************
\subsection*{Conference abstracts}

\begin{enumerate}

\item \textbf{Wilkinson, R.D.} and Lichtwark, G.A., Measuring A Rider's Centre Of Mass Displacement During Non-seated Cycling Using A Single Inertial Measurement Unit. \textit{Annual Meeting of the American-College-of-Sports-Medicine (ACSM)}, San Francisco, CA, United States, 26-30 May 2020.

\item \textbf{Wilkinson, R.D.}, Lichtwark, G.A., and Cresswell, A.G., Effect of lateral bicycle dynamics on rider biomechanics during non-seated cycling. \textit{XXVII Congress of the International-Society-of-Biomechanics (ISB)}, Calgary, ON, Canada, 31 July - 4 August 2019.

\item \textbf{Wilkinson, R.D.}, Williams, J., Marcus, M., and Carver, T., Effect of chamois design on rider comfort and saddle pressure during sub-maximal cycling. \textit{Annual Meeting of the American-College-of-Sports-Medicine (ACSM)}, Orlando, FL, United States, 28 May - 1 June 2019.

\item \textbf{Wilkinson, R.D.}, Lichtwark, G.A., and Cresswell, A.G., Mechanical energetics relating to rider centre of mass motion during non-seated cycling. \textit{42nd Annual Meeting of the American-Society-of-Biomechanics (ASB)}, Rochester, MN, United States, 8-11 August 2018.

\item \textbf{Wilkinson, R.D.}, Lichtwark, G.A. and Cresswell, A.G., Stand and deliver: Muscle activity and mechanical energetics of the lower limb during cycling. \textit{Annual Meeting of the American-College-of-Sports-Medicine (ACSM)}, Minneapolis, MN, United States, 29 May - 2 June 2018.

\item \textbf{Wilkinson, R.D.}, Lichtwark, G.A., and Cresswell, A.G., Sit, stand, rock and roll: Towards optimal pedalling posture. \textit{Human-Movement-and-Nutrition-Sciences (HMNS) Post-Graduate Conference}, Brisbane, QLD, Australia, October 2018.

\item \textbf{Wilkinson, R.D.}, Lichtwark, G.A., and Cresswell, A.G., Sit v stand: Effect of cycling position on lower limb mechanics. \textit{Human-Movement-and-Nutrition-Sciences (HMNS) Post-Graduate Conference}, Brisbane, QLD, Australia, October 2017.

\item \textbf{Wilkinson, R.D.}, Lichtwark, G.A., and Cresswell, A.G., Effect of bike sway and rider position on mechanical effectiveness and efficiency in cycling. \textit{Human-Movement-and-Nutrition-Sciences (HMNS) Post-Graduate Conference}, Brisbane, QLD, Australia, October 2016.

\end{enumerate}

% ***************************************************
\section*{Contributions by others to the thesis}

\begin{instructional}
	List the significant and substantial inputs made by others to the research, work and writing represented and/or reported in the thesis. These could include significant contributions to: the conception and design of the project; non-routine technical work; analysis and interpretation of research data; drafting significant parts of the work or critically revising it so as to contribute to the interpretation. 

	If no one contributed significantly then state ``No contributions by others''.
\end{instructional}

Dr Glen A. Lichtwark and Dr Andrew G. Cresswell significantly contributed to the initial concept and ongoing design of the research project, the analysis and interpretation of experimental data, and the critical revision of written work.

% ***************************************************
\section*{Statement of parts of the thesis submitted to qualify for the award of another degree}

No works submitted towards another degree have been included in this thesis.

% ***************************************************
\section*{Research involving human or animal subjects}

Ethical approval was obtained from the School of Human Movement and Nutrition Sciences Ethics Committee at The University of Queensland (Approval Number(s): HMS16/1409, HMS17/0908) and The University of Queensland Institutional Human Research Ethics Committee (Approval Number(s): 2020000354). All methods were performed in accordance with the relevant ethical guidelines. Prior to data collection, written informed consent was obtained from all participants. Ethical approval forms can be found in Appendix \ref{Chap:F}.

\clearpage

% ***************************************************
%\section*{Acknowledgments}

%\clearpage

% ***************************************************
\section*{Financial support}

Ross D. Wilkinson was supported by an Australian Government Research Training Program Scholarship.

% ***************************************************
\section*{Keywords}

\begin{instructional}
	Maximum 10 words; use lower case throughout, separating words/phrases with commas. For example:
\end{instructional}

Cycling, biomechanics, non-seated, bicycle lean, standing, centre of mass, pedaling mechanics, joint power

\section*{Australian and New Zealand Standard Research Classifications (ANZSRC)}

\begin{instructional}
Provide data that links your thesis to the disciplines and discipline clusters in the Federal Government’s Excellence in Research for Australia (ERA) initiative.

Please allocate the thesis a \textbf{maximum of 3} \href{http://www.abs.gov.au/Ausstats/abs@.nsf/Latestproducts/6BB427AB9696C225CA2574180004463E?opendocument}{Australian and New Zealand Standard Research Classifications (ANZSRC) codes} at the \textbf{6 digit level} and include the descriptor and a percent weighting for each code. Total percent must add to 100.

Example:
\end{instructional}

\noindent
ANZSRC code: 110601, Biomechanics, 90$\%$ \\
ANZSRC code: 110602, Exercise Physiology, 5$\%$ \\
ANZSRC code: 110603, Motor Control, 5$\%$

% ***************************************************
\section*{Fields of Research (FoR) Classification}

\begin{instructional}
Allows for categorisation of the thesis according to the field of research. 

Please allocate the thesis a \textbf{maximum of 3} \href{http://www.abs.gov.au/Ausstats/abs@.nsf/Latestproducts/6BB427AB9696C225CA2574180004463E?opendocument}{Fields of Research (FoR) Codes} at the \textbf{4 digit level} and include the descriptor and a percent weighting for each code. Total percent must add to 100. 

Example:
\end{instructional}

FoR code: 1106, Human Movement and Sports Sciences, 100$\%$

\clearpage

%If you wish to add a dedication (if appropriate), do so here.
%If not, comment out from here...
	\rmfamily
	\normalfont

	\begin{vplace}[1]
		\begin{center}
		    This work is dedicated to my parents, who sacrificed all they had for the sake of my sporting interests and education. To my family and friends - thank you for your love, endless support, and encouragement. To my mentors and colleagues - I wouldn't have made it this far without your wisdom, generosity, compassion, and patience. I hope this work stands as a testament to the belief that we can learn anything if we persevere.
		\end{center}
		
	\end{vplace}

%... to here.

\clearpage
\pagestyle{headings}

%These generate the table of contents, list of figures, and list of tables from items tagged with a \label{} command.
\tableofcontents
	\clearpage
\listoffigures
	\clearpage
\listoftables

\input{symbols.tex}

% *************** End of front matter ***************